---
title: "Project"
author: "Nicholas Adams,Waleed AlAmri,Rachel Ferrara, Megan Campbell,Angela Andrews"
date: "4/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
```

```{r}
conStudent <- src_mysql(
                   host = "ba-isdsclass-programdev.lsu.edu",
                  port = 3306,
                  user = 'student',
                  password = 'student',
                  dbname = "isds_3105")
conStudent
```

```{r}
college <- tbl(conStudent, 'college')
continuingeducation <- tbl(conStudent,'continuingeducation')
currentposition <- tbl(conStudent,'currentposition')
employment <- tbl(conStudent,'employment')
fulltime <- tbl(conStudent,'fulltime')
graduation <- tbl(conStudent,'graduation')
graduationtitle <- tbl(conStudent,'graduationtitle')
internship <- tbl(conStudent,'internship')
location <- tbl(conStudent,'location')
organization <- tbl(conStudent,'organization')
paidinternship <- tbl(conStudent,'paidinternship')
parentcompany <- tbl(conStudent,'parentcompany')
title <- tbl(conStudent,'title')
volunteering <- tbl(conStudent,'volunteering')
college <- as_data_frame(college)
continuingeducation <- as_data_frame(continuingeducation)
currentposition <- as_data_frame(currentposition)
employment <- as_data_frame(employment)
fulltime <- as_data_frame(fulltime)
graduation <- as_data_frame(graduation)
graduationtitle <- as_data_frame(graduationtitle)
internship <- as_data_frame(internship) 
location <- as_data_frame(location) 
organization <- as_data_frame(organization) 
paidinternship <- as_data_frame(paidinternship) 
parentcompany <- as_data_frame(parentcompany) 
title <- as_data_frame(title) 
volunteering <- as_data_frame(volunteering) 
```

##GPA by College

GPA is used to test how well each college's staff is preparing and teaching their students. 

```{r}
editedGraduationUnder <- graduation %>% filter(degreeLevel == 'Bachelor') %>%  select(graduationId,idCollege,cumulativeGPA)
averageGPAcollegeUnder <- editedGraduationUnder %>% group_by(idCollege) %>% summarise(avg = mean(cumulativeGPA, na.rm = T))
collegeGPAdataUnder <- right_join(averageGPAcollegeUnder, college, by = "idCollege", copy = T) %>% filter(idCollege != 'L') %>% filter(idCollege != 'NA') %>% filter(idCollege != 'MCOM') %>% filter(idCollege != 'VETM')

collegeGPAdataUnder%>% arrange(desc(avg)) %>% select(avg, collegeName) %>% kable()

ggplot(data = collegeGPAdataUnder)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('GPA by College(Undergraduate)') 
```

```{r}
editedGraduationMaster <- graduation %>% filter(degreeLevel == 'Master') %>%  select(graduationId,idCollege,cumulativeGPA)
averageGPAcollegeMaster <- editedGraduationMaster %>% group_by(idCollege) %>% summarise(avg = mean(cumulativeGPA, na.rm = T))
collegeGPAdataMaster <- right_join(averageGPAcollegeMaster, college, by = "idCollege", copy = T) %>% filter(idCollege != 'NA') %>% filter(idCollege != 'L')

collegeGPAdataMaster%>% arrange(desc(avg)) %>% select(avg, collegeName) %>% kable()

ggplot(data = collegeGPAdataMaster)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('GPA by College(Masters)') 
```

```{r}
editedGraduationPhD <- graduation %>% filter(degreeLevel == 'PhD') %>%  select(graduationId,idCollege,cumulativeGPA)
averageGPAcollegePhD <- editedGraduationPhD %>% group_by(idCollege) %>% summarise(avg = mean(cumulativeGPA, na.rm = T))
collegeGPAdataPhD <- right_join(averageGPAcollegePhD, college, by = "idCollege", copy = T) %>% filter(idCollege != 'NA') %>% filter(idCollege != 'ADSN') %>% filter(idCollege != 'L')

collegeGPAdataPhD%>% arrange(desc(avg)) %>% select(avg, collegeName) %>% kable()

ggplot(data = collegeGPAdataPhD)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('GPA by College(PhD)') 
```

##Number of Graduates by College

The number of graduates from each college was counted to consider how well the colleges are producing graduates.

```{r}
numcollege <- left_join(graduation, college, by = 'idCollege', copy = T) %>% select(graduationId, collegeName) %>% group_by(collegeName) %>% count()
numcollegeUnder <- left_join(graduation, college, by = 'idCollege', copy = T) %>% filter(degreeLevel == 'Bachelor') %>% select(graduationId, collegeName) %>% group_by(collegeName) %>% count()

numcollegeUnder%>% arrange(desc(n)) %>% select(n, collegeName) %>% kable()

ggplot(data = numcollegeUnder)+
  geom_col(aes(x = collegeName, y = n, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Number of Graduates(Undergraduate)')
```

```{r}
numcollegeMaster <- left_join(graduation, college, by = 'idCollege', copy = T) %>% filter(degreeLevel == 'Master') %>% select(graduationId, collegeName) %>% group_by(collegeName) %>% count()

numcollegeMaster%>% arrange(desc(n)) %>% select(n, collegeName) %>% kable()

ggplot(data = numcollegeMaster)+
  geom_col(aes(x = collegeName, y = n, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Number of Graduates(Masters)')
```

```{r}
numcollegePhD <- left_join(graduation, college, by = 'idCollege', copy = T) %>% filter(degreeLevel == 'PhD') %>% select(graduationId, collegeName) %>% group_by(collegeName) %>% count()

numcollegePhD%>% arrange(desc(n)) %>% select(n, collegeName) %>% kable()

ggplot(data = numcollegePhD)+
  geom_col(aes(x = collegeName, y = n, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Number of Graduates(PhD)')
```

```{r}
gradandcollege <- left_join(graduation, college, by = 'idCollege', copy = T)
```

```{r}
fulltime <- fulltime %>% mutate(graduationId = idCurrentPosition)
```

##Average Salary by College

Average salary was compared to see which college is producing the most valuable students to companies. 

```{r}
saldataUnder <- gradandcollege %>% left_join(fulltime) %>% filter(degreeLevel == 'Bachelor') %>% select(graduationId, collegeName, salary) %>% filter(salary <= 1500000) %>% filter(salary >= 15080) %>% group_by(collegeName) %>% summarise(avg = mean(salary, na.rm = T)) %>% filter(collegeName != 'NA') 

saldataUnder %>% arrange(desc(avg)) %>% kable()

ggplot(data = saldataUnder)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Salaries by College(Undergraduate)') 
```

```{r}
saldataMaster <- gradandcollege %>% left_join(fulltime) %>% filter(degreeLevel == 'Master') %>% select(graduationId, collegeName, salary) %>% filter(salary <= 200000) %>% filter(salary >= 15080)%>% group_by(collegeName) %>% summarise(avg = mean(salary, na.rm = T)) %>% filter(collegeName != 'NA') 

saldataMaster %>% arrange(desc(avg)) %>% kable()

ggplot(data = saldataMaster)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Salaries by College(Master)') 
```

```{r}
saldataPhD <- gradandcollege %>% left_join(fulltime) %>% filter(degreeLevel == 'PhD') %>% select(graduationId, collegeName, salary) %>% filter(salary <= 200000) %>% filter(salary >= 15080)%>% group_by(collegeName) %>% summarise(avg = mean(salary, na.rm = T)) %>% filter(collegeName != 'NA') 

saldataPhD %>% arrange(desc(avg)) %>% kable()

ggplot(data = saldataPhD)+
  geom_col(aes(x = collegeName, y = avg, fill = collegeName))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Salaries by College(PhD)') 
```

```{r}
bacheloractivity <- gradandcollege %>%filter(degreeLevel == 'Bachelor') 

Underactivity <- bacheloractivity %>% select(graduationId,graduationExpectedActivity,collegeName)%>% mutate(employed = ifelse(bacheloractivity$graduationExpectedActivity == c('Employed full-time','Employed part-time',"Employed full-time (on average 30 hours or more per week)","Employed part-time (on average less than 30 hours per week)"),yes = 'Employed', no = 'Not Employed')) %>% mutate(education = ifelse(bacheloractivity$graduationExpectedActivity == c("Graduate or professional school","Continuing education", "Enrolled in a program of continuing education"), yes = 'Continuing Education', no = 'Not Continuing Education'))%>% filter(!is.na(graduationExpectedActivity),!is.na(employed))%>%filter(graduationExpectedActivity != 'Other') %>% filter(graduationExpectedActivity != 'NA') %>% filter(graduationExpectedActivity != 'Starting or raising a family') %>%filter(graduationExpectedActivity != 'Volunteer activity') %>% filter(graduationExpectedActivity != 'Military service') %>% filter(graduationExpectedActivity != 'Seeking employment') %>% filter(graduationExpectedActivity != 'Volounteer activity') %>% filter(graduationExpectedActivity != 'Not seeking employment') %>% filter(graduationExpectedActivity != 'Not seeking employment or continuing education at that time') %>% filter(graduationExpectedActivity != 'Participating in a volunteer or service program') %>% filter(graduationExpectedActivity != 'Serving in the U.S. military') %>% filter(graduationExpectedActivity != 'Planning to continue education')%>% filter(graduationExpectedActivity != 'Planning to continue education but not yet enrolled')
```

```{r}
MSactivity <- gradandcollege %>%filter(degreeLevel == 'Master')

Masteractivity <- MSactivity %>% select(graduationId,graduationExpectedActivity,collegeName) %>% mutate(employed = ifelse(MSactivity$graduationExpectedActivity == c('Employed full-time','Employed part-time',"Employed full-time (on average 30 hours or more per week)","Employed part-time (on average less than 30 hours per week)"),yes = 'Employed', no = 'Not Employed')) %>% mutate(education = ifelse(MSactivity$graduationExpectedActivity == c("Graduate or professional school","Continuing education", "Enrolled in a program of continuing education"), yes = 'Continuing Education', no = 'Not Continuing Education'))%>% filter(!is.na(graduationExpectedActivity),!is.na(employed))%>%filter(graduationExpectedActivity != 'Other') %>% filter(graduationExpectedActivity != 'NA') %>% filter(graduationExpectedActivity != 'Starting or raising a family') %>%filter(graduationExpectedActivity != 'Volunteer activity') %>% filter(graduationExpectedActivity != 'Military service') %>% filter(graduationExpectedActivity != 'Seeking employment') %>% filter(graduationExpectedActivity != 'Volounteer activity') %>% filter(graduationExpectedActivity != 'Not seeking employment') %>% filter(graduationExpectedActivity != 'Not seeking employment or continuing education at that time') %>% filter(graduationExpectedActivity != 'Participating in a volunteer or service program') %>% filter(graduationExpectedActivity != 'Serving in the U.S. military') %>% filter(graduationExpectedActivity != 'Planning to continue education')%>% filter(graduationExpectedActivity != 'Planning to continue education but not yet enrolled')

kable(head(Masteractivity))

```

```{r}

Dactivity <- gradandcollege %>%filter(degreeLevel == 'PhD')

PhDactivity <- Dactivity %>% select(graduationId,graduationExpectedActivity,collegeName) %>% mutate(employed = ifelse(Dactivity$graduationExpectedActivity == c('Employed full-time','Employed part-time',"Employed full-time (on average 30 hours or more per week)","Employed part-time (on average less than 30 hours per week)"),yes = 'Employed', no = 'Not Employed')) %>% mutate(education = ifelse(Dactivity$graduationExpectedActivity == c("Graduate or professional school","Continuing education", "Enrolled in a program of continuing education"), yes = 'Continuing Education', no = 'Not Continuing Education'))%>% filter(!is.na(graduationExpectedActivity),!is.na(employed))%>%filter(graduationExpectedActivity != 'Other') %>% filter(graduationExpectedActivity != 'NA') %>% filter(graduationExpectedActivity != 'Starting or raising a family') %>%filter(graduationExpectedActivity != 'Volunteer activity') %>% filter(graduationExpectedActivity != 'Military service') %>% filter(graduationExpectedActivity != 'Seeking employment') %>% filter(graduationExpectedActivity != 'Volounteer activity') %>% filter(graduationExpectedActivity != 'Not seeking employment') %>% filter(graduationExpectedActivity != 'Not seeking employment or continuing education at that time') %>% filter(graduationExpectedActivity != 'Participating in a volunteer or service program') %>% filter(graduationExpectedActivity != 'Serving in the U.S. military') %>% filter(graduationExpectedActivity != 'Planning to continue education')%>% filter(graduationExpectedActivity != 'Planning to continue education but not yet enrolled')

kable(head(PhDactivity))

```

##Percent of Graduating Class Continuing Education or Employed

The percent of graduating class that is either continuing education or employed was considered to see how well each college places their student in the next step of their lives.

```{r}
Undereducation <- Underactivity %>% group_by(collegeName) %>% count(education)

Undereducationdata <- Undereducation%>% spread(education, "n")

Underpereducation <- Undereducationdata %>% summarise(perc = `Continuing Education`/`Not Continuing Education`*100)

Underpereducation %>% arrange(desc(perc)) %>% kable()

ggplot(data = Underpereducation)+
  geom_col(aes(x = collegeName, y = perc, fill = collegeName))+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Continuing Education by College(Undergraduate)')
```

```{r}
Mastereducation <- Masteractivity %>% group_by(collegeName) %>% count(education)

Mastereducationdata <- Mastereducation%>% spread(education, "n")

Masterpereducation <- Mastereducationdata %>% summarise(perc = `Continuing Education`/`Not Continuing Education`*100)

Masterpereducation %>% arrange(desc(perc)) %>% kable()

ggplot(data = Masterpereducation)+
  geom_col(aes(x = collegeName, y = perc, fill = collegeName))+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Continuing Education by College(Master)')
```

```{r}
Underemployeddata <- Underactivity %>% group_by(collegeName) %>% count(employed)

Underrealemployeddata <- Underemployeddata%>% spread(employed, "n")

Underpercemployed <- Underrealemployeddata %>% summarise(perc = Employed/`Not Employed`*100)

Underpercemployed %>% arrange(desc(perc)) %>% kable()

ggplot(data = Underpercemployed)+
  geom_col(aes(x = collegeName, y = perc, fill = collegeName))+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Employment by College (Undergraduate)')
```

```{r}
Masteremployeddata <- Masteractivity %>% group_by(collegeName) %>% count(employed)

Masterrealemployeddata <- Masteremployeddata%>% spread(employed, "n")

Masterpercemployed <- Masterrealemployeddata %>% summarise(perc = Employed/`Not Employed`*100)

Masterpercemployed %>% arrange(desc(perc)) %>% kable()

ggplot(data = Masterpercemployed)+
  geom_col(aes(x = collegeName, y = perc, fill = collegeName))+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Employment by College (Master)')
```

```{r}
PhDemployeddata <- PhDactivity %>% group_by(collegeName) %>% count(employed)

PhDrealemployeddata <- PhDemployeddata%>% spread(employed, "n")

PhDpercemployed <- PhDrealemployeddata %>% summarise(perc = Employed/`Not Employed`*100)

PhDpercemployed %>% arrange(desc(perc)) %>% kable()

ggplot(data = PhDpercemployed)+
  geom_col(aes(x = collegeName, y = perc, fill = collegeName))+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ggtitle('Employment by College(PhD)')
```